Multimodal Information Extraction of Supermarket Leaflets

### 2025-02-22 paper/main_ilyi.tex
Pre-final paper version of ilyi.

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

% This is also not strictly necessary, and may be commented out.
% However, it will improve the aesthetics of text in
% the typewriter font.
\usepackage{inconsolata}

%Including images in your LaTeX document requires adding
%additional package(s)
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{hyperref}

@@

\section{Introduction}
\subsection{Motivation}
\subsection{Problem Statement}
\subsection{Contributions}

\section{Related Works}
\indent{\textbf{Deal Detection}.}
\indent{\textbf{Optical Character Recognition}.}
OCR-Model-Driven methods use OCR tools to acquire text
and bounding box information. Subsequently, they rely
on the models to integrate text, layout, and visual data.

\indent{\textbf{Information Extraction}.}

\section{Deal Detection}
    \subsection{Datasets}
    \subsection{Model}
    \subsection{Experiments}

\section{Information Extraction of Supermarket Deals}
    % - Subsequent to detection and separation of deals, the goal is to extract useful information from it.
    % - Difficulties
    %     - High variety of layouts, colors, fonts, sizes
    %         - Deal sometimes have superscripted decimals and no separation dot.
    %         - Original prices are sometimes striked through.
    %         - Deal prices are sometimes in a very different font to the ones it was trained on.
    %         - Overlapping elements
    %     - Multimodal information
    %         - Product image: visual
    %         - Text: alphabetical + numerical
    %     - OCR errors
    %     - Eigennamen (mostly brand names)
    %     - German language
    % - Goal: Extract 
    %     - Product name
    %     - brand
    %     - original price
    %     - deal price
    %     - unit
Subsequently to the detection and separation of supermarket deals, the overarching goal of using these in applications requires the extraction of the various useful information from these deals. 
The task of extracting meaningful and structured information from supermarket deal images is a complex and multi-faceted problem, requiring the integration of various modalities, overcoming challenges posed by noisy data, and leveraging deep learning techniques. This section provides a detailed exploration of the underlying challenges, methodologies, and the theoretical and practical framework needed to perform such extraction.

\subsection{Challenges in Information Extraction for Supermarket Deals}

Supermarket deal extraction involves a series of challenges that require careful attention and sophisticated methods to resolve. These challenges include the diversity of layout and format in the images, the multimodal nature of the data, the frequent occurrence of Optical Character Recognition (OCR) errors, and the specific linguistic and cultural issues posed by the German language.

% \subsubsection{High Variety in Layouts and Visual Elements}
\indent{\textbf{High Variety in Layouts and Visual Elements.}}
Supermarket deals exhibit considerable variability in layout, color schemes, font choices, and text sizes, which hinders automatic detection and extraction. Deals are often printed with superscripted decimals or without clear separation between integer and fractional components of the price (e.g., "2.29" might appear as "229"). This phenomenon exacerbates OCR difficulties. Additionally, original prices may be struck through, while deal prices are sometimes printed in drastically different fonts than those the OCR model was trained on. Furthermore, overlapping elements—such as product images or additional textual information—often interfere with text extraction and localization. These diverse visual elements require robust and flexible models capable of accommodating such variability.

% \subsubsection{Multimodal Information}
\indent{\textbf{Multimodal Information.}}
Supermarket deal images consist of both visual (e.g., product images) and textual (e.g., brand names, prices, and product descriptions) modalities, each contributing different types of information. While text provides rich information about the product’s identity, pricing, and units, the image serves as a supplementary modality that aids in product identification, brand recognition, and other visual cues. Multimodal fusion, where information from both text and image domains is integrated, is a key challenge, and existing models must effectively leverage both sources to provide high-quality extraction.

\indent{\textbf{Conversion Errors and Linguistic Challenges.}}
OCR systems frequently introduce errors, particularly when dealing with non-standard fonts, noisy backgrounds, or small text. This is particularly problematic when extracting numerical data such as prices. Common errors include misinterpretation of decimals and digits (e.g., "229" instead of "2.29") and missing characters. To address this, OCR post-processing steps, such as error correction using context-based reasoning or dedicated error models, are required. Additionally, the presence of proper nouns—particularly brand names—adds complexity, as these entities may not appear in typical language models, making their extraction difficult. Furthermore, the German language presents its own set of challenges, including compound words, specific punctuation conventions, and diverse word forms, all of which must be accounted for in any robust extraction model.

\indent{\textbf{Ambiguities.}}
The presence of overlapping elements—whether textual or graphical—can lead to a degradation in the extraction accuracy. This challenge is compounded when the overlap involves essential information, such as when the original price is partially obscured by a product image or strike-through marks. Furthermore, ambiguity in labeling and the context in which different pieces of information appear in the deal image can create difficulties in associating text with the correct object (e.g., associating a price with a product rather than the surrounding descriptive text).

\subsection{Formal Approach}
Formally, the task of information extraction from a deal image can be defined as the identification of specific entities, including the product name, brand, original price, deal price, and unit. We define the following set of output labels:
\begin{equation}
\mathcal{Y} = \{y_{i}\}_{i=1}^{n},
\end{equation}
where \(y_{i}\) represents the $i$-th entity in the deal image. In the following, the entities are defined as:
\begin{itemize}
    \item \(y_{\text{product\_name}}\): The name of the product.
    \item \(y_{\text{brand}}\): The brand of the product.
    \item \(y_{\text{original\_price}}\): The original price of the product.
    \item \(y_{\text{deal\_price}}\): The deal price of the product.
    \item \(y_{\text{unit}}\): The unit of the product (e.g., weight, volume).
\end{itemize}

To extract these entities, one aims to learn a function \(f(\cdot)\) that maps an input image \(I\) to the desired outputs:
\begin{equation}
    f: I \to \mathcal{Y},
\end{equation}
where \( I \) is the deal image, and \( \mathcal{Y} \) represents the extracted entities.

The choice of the function \( f(\cdot) \) is an architectural decision that depends on the specific requirements of the task, the nature of the data, the available resources, and the desired performance metrics.

\subsection{Architectural Approaches to Information Extraction}
- Classical methods: Rule-based, template-based
- Traditional methods: Dividing into multi-stage
    - 1. Text detection
    - 2. Text recognition
    - 3. Key-value pair extraction
- OCR + LLM
- End2End: Vision Encoder Decoder Model
    - Donut 
- E2E: LVLM

\subsection{Dataset Creation}
Since there was no dataset to evaluate or train, a custom dataset was created by manually annotating a collection of supermarket deal images. The dataset includes images with diverse layouts, fonts, and colors to ensure robustness. Each image is annotated with bounding boxes for text regions and corresponding labels for product name, brand, original price, deal price, and unit. The dataset is split into training, validation, and test sets to facilitate model development and evaluation. 
% Table \ref{tab:dataset} provides an overview of the custom dataset and Figure \ref{fig:dataset} shows sample images with annotations.
% TODO: Add table and figure

\subsection{Optical Character Recognition}
Optical Character Recognition (OCR) is a critical component in the information extraction pipeline. Several OCR solutions were evaluated to determine the most suitable one for our task. The evaluated OCR tools include Tesseract, EasyOCR, PaddleOCR, and DocTR. Each tool was assessed based on its accuracy, speed, and ability to handle diverse fonts and layouts.

\subsection{Experiments}
\indent{\textbf{Setup}.}
The experiments were conducted to compare the performance of different OCR solutions and end-to-end models for information extraction. The evaluation metrics include precision, recall, and F1-score for each entity type (product name, brand, original price, deal price, and unit). The experiments were performed on the test set of the custom dataset.

\indent{\textbf{Results}.}
The results of the experiments indicate that the end-to-end models, such as the Vision Encoder Decoder Model (e.g., Donut) and Large Vision-Language Models (LVLM), outperform traditional OCR-based methods in terms of accuracy and robustness. The OCR + LLM approach also shows promising results, particularly in handling OCR errors and extracting key-value pairs.

\subsection{Evaluation}
The evaluation of the models was conducted using standard metrics, including precision, recall, and F1-score. The comparison between LVLM, OCR + LLM, and Donut models highlights the strengths and weaknesses of each approach. The results demonstrate that end-to-end models provide superior performance in extracting structured information from supermarket deal images.

\section{Application}
\subsection{Database}

\subsection{Optical Character Recognition}

    \subsection{Experiments}
    \indent{\textbf{Setup}.}
    - Common OCR solutions: Tesseract, EasyOCR, PaddleOCR, DocTR
    - Goal: Which OCR solution is the best for our task?
    -

    \subsection{Evaluation}
    - LVLM, OCR+LLM, Donut
    - OCR + LLM Vergleich

>>>>

\usepackage{microtype}

\usepackage{inconsolata}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{subfig}

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{verbatim}

@@

\section{Introduction}

\subsection{Motivation}
When one thinks of a paper-heavy country, Germany is likely to come to mind. The country's affinity for printed materials is deeply ingrained in its culture, with activities such as browsing through supermarket leaflets serving as a familiar ritual for many individuals. These leaflets have long been an essential medium for consumers to plan their purchases, discover special offers, and compare prices. The tactile nature of printed leaflets, the excitement of spotting discounts, and the satisfaction of making a well-informed shopping decision contribute to their enduring appeal.
However, the rapid advancement of digital technology is transforming consumer behavior. While some individuals still find browsing through physical leaflets a nostalgic and even relaxing experience, younger generations are increasingly turning to digital alternatives for their shopping needs. The modern shopper expects convenience, efficiency, and instant access to information, yet the digital transformation of supermarket leaflets lags significantly behind other aspects of e-commerce and retail technology.

Despite the existence of applications that compile digital versions of these leaflets, their usability and overall quality remain suboptimal. Many digital leaflets are plain documents, failing to leverage the full potential of interactivity, searchability, and intelligent recommendation systems. There is a pressing need for sophisticated digital solutions that not only present supermarket deals but also enable consumers to interact with them in a meaningful way.

Possibly, one can imagine a seamless digital platform where users can search for specific items across multiple supermarket chains, apply dietary filters to instantly highlight suitable products, or receive personalized recommendations based on past preferences. A tool that allows consumers to create a dynamic shopping list, track price trends over time, and receive real-time updates on the best deals would revolutionize the way people approach grocery shopping. Yet, these features remain largely absent from existing digital leaflet solutions. Furthermore, while price comparison websites and applications have become commonplace for electronics, fashion, and travel, the grocery sector remains an overlooked frontier. Consumers are often left navigating multiple supermarket websites or manually cross-referencing prices to find the best deals—a tedious and inefficient process. A comprehensive digital ecosystem that integrates grocery price comparisons, personalized discounts, and AI-driven shopping assistants could bridge this gap and act as a workaround, offering an enhanced, data-driven shopping experience tailored to individual needs.

\subsection{Problem Statement}
The primary objective of this research is to develop an intelligent system for extracting, structuring, and utilizing information from supermarket leaflets in a digital format. Given a set of supermarket leaflets $\mathbf{L} = \{L_1, L_2, ..., L_n\}$, where each $L_i$ contains a set ofproduct deals $\{p_j\}_{j=1}^{m} \in P_{L_i}$, the goal is to extract and normalize these attributes to enable efficient querying and comparison across multiple supermarket chains.

Each product deal $p_j$ is represented by a set of attributes $\{y_1, y_2, ..., y_k\}$, where each attribute $y_k$ represents a specific entity. Common attributes include the name and brand of the product, the original and deal prices, the unit (e.g., weight, volume), and the product image. The challenge lies in detecting and separating deals from the leaflet, extracting structured information from the detected deals, normalizing the extracted information to ensure consistency and comparability, and validating the extracted information to ensure accuracy.

To enable advanced applications such as the abovementioned, one has to tackle different challenges, such as:
\begin{enumerate}
    \item Detecting and separating deals $p$ from the leaflet $L$.
    \item Extracting structured information from the detected deals.
    \item Normalizing the extracted information to ensure consistency and comparability.
    \item Validating the extracted information to ensure accuracy.
\end{enumerate}

Therefore, a function $f: \mathcal{L} \to \mathcal{D}$ has to be designed, that maps a set of leaflets $\mathcal{L}$ to a set of structured deals $\mathcal{D}$, where each deal $d \in \mathcal{D}$ is represented by a set of attributes $\{y_1, y_2, ..., y_k\}$.

\subsection{Contributions}
To build this mapping, this work presents the conducted research and the practical result of a comprehensive approach to digitalizing and structuring supermarket leaflet information by leveraging advanced computer vision, OCR, and language models, as shown in \figref{fig:overview}. A \emph{modular pipeline} is proposed that encompasses deal detection, information extraction and normalization to transform raw leaflet data into structured, actionable insights.

\begin{figure*}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/overview.png}
    \caption{Overview of the proposed approach to digitalizing and structuring supermarket leaflet information.}
    \label{fig:overview}
\end{figure*}

To enhance usability, an \emph{interactive application} is developed that allows consumers to browse, filter, and compare deals across multiple supermarket chains. It provides essential search, tracking, and comparison features for improved deal discovery.

To facilitate research and development in this domain, \emph{Leaflet-IE}, an end-to-end benchmark dataset for information extraction from supermarket leaflets, is introduced. This dataset stands as a valuable resource for evaluating and comparing different models and approaches in the field.

\subsection{Structure}
This work is structured to systematically address supermarket deal detection and information extraction, beginning with a scientific overview of related works in Section 2, which categorizes existing approaches and identifies their limitations. Section 3 provides a detailed analysis of the nature of supermarket leaflet data, formalizing challenges such as layout variability, OCR noise, and multimodal fusion. Section 4 focuses on deal detection, covering dataset creation, model selection, and experimental results for identifying promotional content. Section 5 extends this to information extraction, detailing dataset construction, model architectures, and performance evaluation for extracting structured product and pricing information. Finally, Section 6 concludes with key contributions, limitations, and future directions, including scalability and temporal reasoning improvements.

\section{Related Works}
\indent{\textbf{Supermarket Leaflet Analysis}.}
The field of supermarket leaflet analysis has evolved to tackle key challenges such as product classification, promotion identification, and customer interaction. Recent studies have leveraged AI-driven techniques to enhance efficiency and accuracy in this domain. For instance, Ladwig et al. \cite{ladwig2023} explore fine-grained product classification in leaflet advertisements, while Arroyo et al. \cite{arroyo2020} employ multi-label classification methods to analyze promotions using both textual and visual information. Additionally, Zhang et al. \cite{zhang2024} introduce PKU-GoodsAD, a dataset designed for unsupervised anomaly detection and segmentation of supermarket goods. Complementing these efforts, Nandkumar et al. \cite{nandkumar2024} propose a multi-level LLM conversational interface to enhance supermarket robot interactions by handling diverse customer intents.

\indent{\textbf{Deal Detection}.}

\indent{\textbf{Information Extraction}.}
Recent advancements in Optical Character Recognition (OCR), Large Language Models (LLMs), and end-to-end multimodal architectures have significantly refined the landscape of document information extraction. These improvements have led to increased accuracy, efficiency, and versatility in processing structured and unstructured text from diverse document formats.

OCR technology has evolved from traditional rule-based approaches to sophisticated deep learning-driven systems. Open-source OCR frameworks such as Tesseract, PaddleOCR, EasyOCR, and DocTR integrate state-of-the-art methodologies to enhance both text detection \cite{liao2019, liao2022} and recognition \cite{shi2015, li2022}. These systems employ deep neural networks to improve accuracy in complex document layouts, including distorted and low-quality images. Preprocessing techniques play a crucial role in optimizing OCR performance. Methods such as rectification \cite{shi2019}, deskewing \cite{pham2022}, and denoising \cite{zhao2018} significantly enhance text clarity by mitigating distortions, reducing noise, and improving alignment. These preprocessing steps are especially useful in real-world scenarios where scanned or photographed documents contain various inconsistencies. LLMs have emerged as powerful tools for refining OCR outputs through context-aware reasoning and semantic corrections. Models like GPT-3 \cite{brown2020} and LLaMA \cite{touvron2023} demonstrate exceptional performance in natural language understanding, enabling improved entity recognition and post-processing of extracted text \cite{yenduri2023}. These models enhance the reliability of OCR by resolving ambiguities, correcting misinterpretations, and providing structured outputs in diverse domains such as finance, healthcare, and legal documentation.

Transformer-based architectures \cite{vaswani2017, dosovitskiy2021} have revolutionized document understanding by integrating vision and language processing into unified frameworks. Models such as Donut \cite{kim2022} and DocFormer \cite{appalaraju2021} leverage attention mechanisms to jointly process textual and visual cues, achieving state-of-the-art performance in document analysis. These models are particularly effective in handling complex layouts, handwriting recognition, and tabular data extraction. Furthermore, multimodal approaches \cite{peng2022, li2021} introduce hybrid learning techniques that enhance information retrieval from scanned documents and digital forms.

The integration of large vision-language models (LVLMs) has further expanded the capabilities of document intelligence. Models such as LLaMA 3.2 Vision \cite{touvron2023}, Qwen-VL \cite{qwen2025}, and GPT-4V \cite{dai2023} utilize large-scale multimodal training to perform sophisticated image-text reasoning. These advancements enable robust document understanding, particularly in scenarios where textual information is closely intertwined with graphical elements. Recent contributions \cite{li2024, wei2024} emphasize the growing potential of LVLMs.

\section{The Nature of Supermarket Leaflet Data}  
Supermarket leaflets constitute a complex multimodal data domain characterized by heterogeneous layouts, ambiguous associations, and noisy signal extraction. As illustrated in \figref{fig:leaflet_data_samples}, these documents present unique computational challenges that demand joint reasoning over visual, textual, and contextual modalities. We formalize the key obstacles below.

Leaflets exhibit extreme variability in design elements that impede automated parsing. The spatial heterogeneity, thus the non-standardized arrangements of product images, prices, and promotions across retailers, is a primary concern. The layout of leaflets can vary significantly, with product images, prices, and promotional text appearing in diverse configurations. Additionally, typography challenges, such as superscripted decimals, struck-through original prices, and font inconsistencies between deals and regular text, further complicate the extraction process. Overlapping elements, where text is occluded by product images or graphical overlays, violate the assumptions of Optical Character Recognition (OCR) systems, which assume planar text.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\linewidth]{figures/samples.png}
    \caption{Common German supermarket deals.}
    \label{fig:leaflet_data_samples}
\end{figure}

Inherently, these leaflets contain both textual and visual information, which must be jointly processed to extract meaningful insights. The textual elements include product names, brands, prices, and units, while the visual components consist of product images, logos, and promotional graphics. The challenge lies in aligning these modalities to extract structured information accurately. Furthermore, the multimodal nature of the data introduces semantic ambiguities, where textual and visual elements maintain non-strict correspondences. For instance, a product image may correspond to multiple text entries, or price interpretation may require visual grounding.

Due to these characteristics, leaflets are prone to linguistic ambiguities, conversion errors and cultural-specific semantics. The German language, with its compound nouns, specific punctuation conventions, and diverse word forms, poses unique challenges for text extraction. OCR errors, such as glyph confusion and tokenization failures, are common due to font artifacts and compound nouns. Multilingual mix, where English promotion terms appear alongside German body text, further complicates the parsing process. Price disambiguation, distinguishing unit prices from total prices without weight references, is a critical challenge. Additionally, more complex offer strategies, such as buy-one-get-one-free, coupon-based discounts, and permanent price reductions, require nuanced interpretation beyond simple text extraction.

Additionally, supermarket leaflets are commonly scheduled for a weekly release, with each leaflet containing a set of deals that are valid for a specific period. However, not every supermarket chain follows this pattern, and some may release leaflets bi-weekly, monthly or even irregularly. This temporal grounding problem requires associating each deal with an expiry time, which is often implicit in the leaflet text. For instance, a deal may be valid only in specific store branches or for a limited time, indicated by small print codes or context clues.

This multifaceted nature demands architectures that jointly optimize for visual-textual alignment, error-resilient parsing, and contextual plausibility. 
\section{Deal Detection}
    \subsection{Datasets}
    \subsection{Model}
    \subsection{Experiments}

\section{Information Extraction of Supermarket Deals}
Subsequently to the detection and separation of supermarket deals, the overarching goal of using these in applications requires the extraction of the various useful information from these deals. 
The task of extracting meaningful and structured information from supermarket deal images is a complex and multi-faceted problem, requiring the integration of various modalities, overcoming challenges posed by noisy data, and leveraging deep learning techniques. This section provides a detailed exploration of methodologies, as well as the theoretical and practical framework needed to perform such extraction.

\subsection{Approaches to Information Extraction}
Formally, the task of information extraction from a deal image can be defined as the identification of specific entities, including the product name, brand, original price, deal price, and unit. We define the following set of output labels:
\begin{equation}
\mathcal{Y} = \{y_{i}\}_{i=1}^{n},
\end{equation}
where $y_{i}$ represents the $i$-th entity in the deal image. In the following, the entities are defined as:
\begin{itemize}
    \item \code{$y_{\text{product\_name}}$}: The name of the product.
    \item \code{$y_{\text{brand}}$}: The brand of the product.
    \item \code{$y_{\text{original\_price}}$}: The original price of the product.
    \item \code{$y_{\text{deal\_price}}$}: The deal price of the product.
    \item \code{$y_{\text{unit}}$}: The unit of the product (e.g., weight, volume).
\end{itemize}

To extract these entities, one aims to learn a function $f(\cdot)$ that maps an input image $I$ to the desired outputs:
\begin{equation}
    f: I \to \mathcal{Y},
\end{equation}
where $ I $ is the deal image, and $ \mathcal{Y} $ represents the extracted entities.

The choice of the function $ f(\cdot) $ is an architectural decision that depends on the specific requirements of the task, the nature of the data, the available resources, and the desired performance metrics.

Among the various existing architectural approaches to learn $ f(\cdot) $ for information extraction, several methodologies have been developed to address the inherent complexities in supermarket deal images. These methods can be categorized into classical approaches, multi-stage traditional models, hybrid OCR-LLM systems, and end-to-end deep learning frameworks. Each paradigm offers distinct advantages, and their applicability depends on computational resources, dataset characteristics, and performance requirements.

\subsubsection{Traditional and Hybrid Approaches for Supermarket Leaflet Information Extraction}  
Information extraction from supermarket leaflets has traditionally been approached through a pipeline of sequential sub-tasks, each addressed by specialized systems. These methods, while extensively studied and methodologically diverse, are typically divided into three main stages: text detection, text recognition, and key-value pair extraction. Although effective in certain scenarios, these approaches suffer from inherent limitations, including computational overhead, error propagation, and suboptimal performance on edge cases. Recent advancements in large language models (LLMs) have led to the development of hybrid OCR-LLM architectures, which combine the precision of optical character recognition (OCR) with the contextual reasoning capabilities of LLMs. Below, we formalize these approaches and discuss their strengths and limitations.

Traditional architectures decompose the problem into independent sub-tasks, each handled by specialized models. While this modularity allows for targeted optimization, it introduces complexity and inefficiencies due to the lack of end-to-end training and error propagation across stages.

The first stage, \emph{text detection}, involves identifying regions in the image $ I \in \mathbb{R}^{H \times W \times 3} $ that contain textual information. This is typically achieved using rule-based systems or neural networks designed for small object detection. The output is a set of $ M $ bounding boxes $ \{b_i\}_{i=1}^{M} $, where each $ b_i = (x_{\text{min}}, y_{\text{min}}, x_{\text{max}}, y_{\text{max}}) $ defines the coordinates of a text region.

Each detected bounding box $ b_i $ is cropped and processed by a \emph{text recognition} model to convert the visual representation into a textual one. This task is challenging due to decoding errors caused by noisy backgrounds, stylized fonts, and ambiguous characters. Given the set of bounding boxes $ \{b_i\}_{i=1}^{M} $, the goal is to predict the corresponding textual entities $ \{t_i\}_{i=1}^{M} $:  
\begin{equation}
\{t_i\}_{i=1}^{M} = \arg\max_{\{t_i\}_{i=1}^{M}} P(\{t_i\}_{i=1}^{M} | \{b_i\}_{i=1}^{M}).
\end{equation}

To use the recognized texts for information extraction, the final stage involves parsing the textual entities into structured outputs. This is typically achieved through rule-based systems or sequence-to-sequence models that generate key-value pairs from the recognized text. The output is a set of structured entities $ \{y_i\}_{i=1}^{N} $, where each $ y_i $ represents a specific entity (e.g., product name, price, unit).

The rise of \emph{large language models }(LLMs) has enabled the development of hybrid OCR-LLM architectures, which combine the strengths of OCR tools with the linguistic and reasoning capabilities of LLMs. These approaches retain the traditional OCR pipeline for converting visual text into machine-readable form but leverage LLMs to enhance robustness through contextual reasoning and structured output generation.

In a hybrid OCR-LLM system, the OCR stage processes the input image $ I $ to produce a set of unstructured text fragments $ T = \{t_i\}_{i=1}^{M} $ with associated spatial coordinates. These fragments are serialized into a prompt and fed into a pretrained LLM, which performs contextual parsing and generates some linguistic output that, in most cases, has to be somewhat postprocessed to extract the desired information.

Hybrid OCR-LLM models offer several advantages over traditional pipelines. By integrating contextual reasoning, they reduce error propagation and improve robustness to noisy or ambiguous inputs. Additionally, the use of pretrained LLMs eliminates the need for task-specific training, simplifying system development. However, these models still rely on the accuracy of the OCR stage, and misrecognitions (e.g., "O" vs. "0") can propagate to the LLM. Furthermore, the computational cost of LLM inference can be prohibitive for large-scale applications.

\subsubsection{End-to-End Models}  
End-to-End (E2E) approaches unify document understanding by learning a direct mapping $ F: I \to \mathcal{D} $ and thus jointly optimize visual feature extraction and semantic parsing. We formalize two dominant paradigms: vision-encoder-decoder architectures and large vision-language models (LVLMs), which share foundational principles but diverge in scale and architectural complexity.

Foundationally, all E2E models decompose into three components: a vision encoder, a representation projection, and a decoder. The vision encoder extracts hierarchical visual features $ \mathbf{Z} = \text{Enc}(I) $, which are then aligned with textual semantics through a representation projection. The decoder generates structured output $ \mathbf{Y} $ autoregressively. \figref{fig:e2e_arch} illustrates the unified architecture.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.5\linewidth]{figures/vlm_arch.png}
    \caption{End-to-End Vision-Encoder-Decoder Architecture.}
    \label{fig:e2e_arch}
\end{figure}

\indent{\textbf{Donut \cite{kim2022}}.} This model employs a compact transformer-based encoder-decoder. Let $ I $ be partitioned into $ N $ patches $ \{\mathbf{p}_i\}_{i=1}^N $, each linearly projected to dimension $ d $.  The used vision encoder $E$ computes an encoded representation mainly through:
\begin{equation}
Z = E(I) = \text{FFN}(\text{MSA}(X) + X) + \text{MSA}(X)
\end{equation}  
where $\text{MSA}$ denotes multi-head self-attention and $\text{FFN}$ a feed-forward network. The hierarchical structure captures local textural patterns as well as global layout cues.

Subsequently, a transformer decoder generates tokens $ Y = (y_1, \dots, y_T) $ using cross-attention over $ Z $:  
\begin{equation}
P(y_t | y_{<t}, Z) = \text{Softmax}(\text{FFN}(\text{MSA}(Y) + Z)),
\end{equation}  
where $ Y $ is the token sequence and $ Z $ the encoded image features. The model is trained end-to-end with a maximum likelihood objective.

\indent{\textbf{Large Vision-Language Models (LVLMs)}.} In the realm of upscaling language models into LLMs, computer vision utilizes this trend to process multimodal data. LVLMs like Llama 3.2 Vision \cite{touvron2023} scale this framework by integrating pretrained vision encoders with large language models (LLMs). These architectures utilize \emph{cross-modal projection}, which aligns visual features with the LLM's embedding space, and \emph{multimodal decoding}, which attends to both visual tokens and textual prompts. The LVLM's strength lies in zero-shot reasoning and leveraging pretrained knowledge for structured output generation.

While smaller models like Donut offer efficiency and direct output generation, LVLMs provide zero-shot reasoning and open-world capabilities at increased computational cost.

\subsection{Dataset Creation}
Since the availability of German supermarket leaflet data is literally not findable, a custom dataset was created by manually annotating a collection of supermarket deal images. The dataset includes images from the most common supermarket chains, each with unique layout, fonts, and colors to, in general, ensure the highest diversity and robustness possible. Each image is annotated with a corresponding label file that includes the desired entities as well as the image identificator. The dataset is split into a training (\~ 80\%) and a validation set (\~ 20\%) for each model to be trained and evaluated on. Table \tabref{tab:ie_dataset} provides an overview of the dataset with the sample size for each entity and \figref{fig:ie_dataset_samples} shows some sample images with their annotations.

\begin{table}[h!]
\centering
\caption{Leaflet-IE Dataset}
\label{tab:ie_dataset}
\begin{tabular}{lc}
\toprule
Entity           & Sample Size \\
\midrule
\code{image\_id}        & 372  \\
\code{brand}            & 357  \\
\code{product\_name}    & 370  \\
\code{original\_price}  & 286  \\
\code{deal\_price}      & 372  \\
\code{weight}           & 369  \\
\bottomrule
\end{tabular}
\end{table}

Even though the Leaflet-IE dataset is relatively small, the reader will be able to see that the dataset is sufficient to evaluate the performance of different IE appraoches as well as to train an competitive end-to-end model. However, as the reader may notice, the dataset is solely focused on single product deals at this point and explicitly does not include

\begin{figure*}[h!]
\centering
\includegraphics[width=0.8\linewidth]{figures/ie_samples.png}
\caption{Sample images with annotations from the custom dataset.}
\label{fig:ie_dataset_samples}
\end{figure*}

\section{Experiments}

The experiments aim to identify the most effective methodology for extracting structured information from supermarket leaflets by evaluating three candidate approaches: (1) Hybrid OCR + LLM pipelines, (2) end-to-end Vision-Encoder-Decoder models, and (3) Large Vision-Language Models (LVLMs). The evaluation is structured into four sequential phases to systematically assess performance and scalability.

First, Plain Deal OCR Performance establishes a baseline by measuring the accuracy of standalone OCR systems in extracting textual content from leaflets, highlighting challenges such as font variability and noisy backgrounds. Second, OCR + LLM Performance evaluates the combined pipeline, where OCR outputs are processed by large language models to resolve ambiguities and extract structured information. Third, Donut Fine-Tuning Performance assesses the Vision-Encoder-Decoder model's effectiveness after fine-tuning on leaflet-specific data, focusing on its ability to handle multimodal inputs without intermediate OCR steps. Finally, the MAIN experiment conducts a comprehensive comparison of Hybrid OCR+LLM, Donut, and LVLM approaches, analyzing their accuracy, robustness, and computational efficiency in real-world leaflet data scenarios.

\subsection{Implementation Details}
The experiments were conducted on two distinct computational workstations - station 1 equipped with an NVIDIA RTX 4080 GPU and station 2 with an NVIDIA GTX 1080 GPU. Therefore, the experiments could be conducted in parallel, ensuring efficient model training and evaluation.

The software stack includegraphics Python with PyTorch\footnote{\href{https://pytorch.org/}{https://pytorch.org/}} as the primary deep learning framework, HuggingFace Transformers \footnote{\href{https://huggingface.co/transformers/}{https://huggingface.co/transformers/}}
for leveraging pre-trained models, ollama \footnote{\href{https://ollama.com/}{https://ollama.com/}}
and lmstudio \footnote{\href{https://lmstudio.ai/}{https://lmstudio.ai/}} for large language models and large vision-language models, and OpenCV \footnote{\href{https://opencv.org/}{https://opencv.org/}} for image preprocessing and data augmentation, as well as various other libraries for data pre-and post-processing, visualization, data management and monitoring \footnote{The codebase is available at \href{https://github.com/ilyii/leaflets}{https://github.com/ilyii/leaflets}}.

% SUB-STUDIES
\subsection{Plain Deal OCR Performance}
The aim of this experiment is to evaluate the performance of various OCR models in extracting text from supermarket leaflets. The evaluation focuses on the accuracies of the extracted entities from \tabref{tab:ie_dataset} and the impact of different normalization levels on the recognition performance. The results provide insights into the robustness and accuracy of different OCR models in handling the complexities of supermarket leaflet data.

The most commonly used OCR models include:
\begin{itemize}
    \item Tesseract \footnote{\href{https://github.com/tesseract-ocr/tesseract}{https://github.com/tesseract-ocr/tesseract}},
    \item EasyOCR \footnote{\href{https://github.com/JaidedAI/EasyOCR}{https://github.com/JaidedAI/EasyOCR}},
    \item PaddleOCR \footnote{\href{https://github.com/PaddlePaddle/PaddleOCR}{https://github.com/PaddlePaddle/PaddleOCR}},
    \item DocTR \footnote{\href{https://github.com/mindee/doctr}{https://github.com/mindee/doctr}}.
\end{itemize}
These models have been trained on a variety of datasets and are known for their robustness and accuracy in text extraction. Since it is likely that none of the OCR models have seen this data before, all 372 samples of the Leaflet-IE dataset were used for evaluation. While OCR alone is not capable of generating key-value mappings, the evaluation used a different set of metrics:

\begin{itemize}
    \item \textbf{Accuracy}: Measures whether the expected value is present in the OCR output.
    \item \textbf{N-gram Accuracy}: Determines if an n-gram substring of the expected value exists in the OCR output, computed for $n \in \left[\frac{|v|}{2}, |v| - 1\right]$, where $|v|$ represents the string length of the ground truth entity value.
\end{itemize}

Given that language-specific OCR errors are common, the evaluation was conducted at different levels of normalization to improve comparability. The text normalization included the following layers:
\begin{itemize}
    \item \textbf{Level 1:} Text stripping and string conversion.
    \item \textbf{Level 2:} Lowercasing and basic replacements.
    \item \textbf{Level 3:} Use-case specific replacements.
    \item \textbf{Level 4:} Punctuation and whitespace removal.
\end{itemize}
Each level includes the normalization steps of the previous levels. Therefore, the normalization procedure can be thought of as a gradual alignment of the OCR output to the ground truth to investigate the OCR model's robustness to different normalization levels. The reader is referred to \secref{app:ocr_normalization} for a detailed description of the normalization procedure.

\indent{\textbf{Results}.} Figure \ref{fig:eval_ocr_accuracies} presents a comparative analysis of various OCR models and normalization levels based on per-entity accuracy. The results highlight several key trends regarding the recognition performance of different entities.

\begin{figure*}
    \begin{tabular}{cc}
      \includegraphics[width=0.5\linewidth]{figures/accuracy_level_1.png} &   \includegraphics[width=0.5\linewidth]{figures/accuracy_level_2.png} \\
    (a) Level 1 & (b) Level 2 \\[6pt]
     \includegraphics[width=0.5\linewidth]{figures/accuracy_level_3.png} &   \includegraphics[width=0.5\linewidth]{figures/accuracy_level_4.png} \\
    (c) Level 3 & (d) Level 4 \\[6pt]
    \end{tabular}
    \caption{Per-Entity Accuracy of OCR Models at Different Normalization Levels.}
    \label{fig:eval_ocr_accuracies}
    \end{figure*}

In general, the OCR models exhibit superior performance in extracting words compared to numerical values. This discrepancy is particularly evident in entity-wise accuracy variations. At the baseline level (normalization level 1), the deal price entity consistently shows the lowest accuracy, whereas the brand entity achieves the highest. The accuracy of other entities varies significantly, with PaddleOCR, EasyOCR, and docTR displaying comparable performance, while Tesseract lags behind.

The overall accuracy improves with increasing normalization levels, with normalization level 4 yielding the highest accuracy gains. At this level, Tesseract continues to struggle with deal price recognition, failing in over 90\% of cases. In contrast, the other OCR models achieve accuracies exceeding 70\% across nearly all entity types.
